Modelling the gas kinematics of an atypical Lyman alpha emitting compact dwarf galaxy

### 2017-06-29 paper/paper.tex
introduction

Observed \lya\ line profiles usually present a single peak shifted
redwards from the line's center. 
Sometimes a double peak is present but the asymmetry persists with 
the peak on the red side being stronger. 
This can be explained as a result of an outflowing shell of neutral Hydrogen.
\citep{2006A&A...460..397V,Orsi12,2012ApJ...751...29Y,2015ApJ...812..123G}.

\tol\ is a compact star forming dwarf galaxy that presents a
strong \lya\ emission with puzzling 
features: the line is highly symmetric, single peaked and broad around the
line's center \citep{Thuan97}.
This kind of profile also raises the question whether some high
redshift LAEs have asymmetric lines because the blue half was
truncated by the intergalactic medium \citep{2007MNRAS.377.1175D}. 
In this case the \lya\ radiation could emerge as a low surface
brightness glow, which may be connected to \lya\ halos, while also
influencing the way LAEs can be used as a probe of reionization
See the review by \cite{2014PASA...31...40D} and the references therein.

@@

The main motivation for the outflowing multiphase model (as presented
in \cite{Gronke2016}) is that some dwarf galaxies are expected to
present outflows.  
Observationally, outflows have been detected in few local dwarf galaxies
\citep{1998ApJ...506..222M,2005MNRAS.358.1453O}. 
Clumpy multiphase outflows are capable to explain \lya\ features around
star-forming galaxies
\citep{2010ApJ...717..289S,2012MNRAS.424.1672D}.

@@

We do not claim that our findings discard the existence of different
undiscovered models that could reproduce the observed features in \tol.
More involved models with multiple kinematic components could
in principple be also sucessful.
We focus on presenting two simple and succesful models that
manage, for the first time, to account for \tol's \lya\ line main
features.

We review first the observational characteristics of
\tol, then we summarize the main features in the multiphase and
rotation models to explain how we fit their free parameters 
to the \tol's \lya\ line shape.
We use the results to interpret them in terms of the galaxy's
dynamical mass and explain why the multiphase model should be prefered
over the rotation model.
We close by discussing the physical scenarios opened up by our
results and the additional observational information needed to clarify
the kinematic nature of \tol.

>>>>

Observed \lya\ line profiles usually present a single peak shifted
redwards from the line's center. 
Sometimes a double peak is present but the asymmetry persists with 
the peak on the red side being stronger. 
This can be explained as the results of multiple \lya\ photon
scatterings through an outflowing shell of neutral Hydrogen
\citep{2006A&A...460..397V,Orsi12,2012ApJ...751...29Y,2015ApJ...812..123G}.

\tol\ is a compact star forming dwarf galaxy that presents a
strong \lya\ emission with puzzling 
features: the line is highly symmetric, single peaked and broad 
\citep{Thuan97}.
This kind of profile also raises the question whether some high
redshift LAEs have asymmetric lines because the blue half was
truncated by the intergalactic medium \citep{2007MNRAS.377.1175D}. 
In this case the \lya\ radiation could emerge as a low surface
brightness glow, which may be connected to \lya\ halos, while also
influencing the way LAEs can be used as a probe of reionization,
see the review by \cite{2014PASA...31...40D} and the references
therein.

@@

The main motivation for the outflowing multiphase model (as presented
in \cite{Gronke2016}) is that some dwarf galaxies are expected to
present outflows.  
Observationally, outflows have been detected in few local dwarf galaxies
\citep{1998ApJ...506..222M,2005MNRAS.358.1453O}. 
Besides, clumpy multiphase outflows are capable to explain \lya\ features around
star-forming galaxies
\citep{2010ApJ...717..289S,2012MNRAS.424.1672D}.

@@

We do not claim that our findings discard the existence of different
undiscovered models that could reproduce the observed features in \tol.
More involved models with multiple kinematic components can
in principle also be sucessful.
We focus on presenting two simple and succesful models that
manage, for the first time, to account for \tol's \lya\ line main
features.

In the next section We review the observational characteristics of
\tol, then we summarize the main features in the multiphase and
rotation models to explain how we fit their free parameters 
to the \tol's \lya\ line shape.
We use the results to interpret them in terms of the galaxy's
dynamical mass and argue why the multiphase model should be prefered
over the rotation model.
We close by discussing the additional observational information needed
to clarify the kinematic nature of \tol.
